What We Eval in the Shadows

### 2021-08-13 paper/main.tex
Capitalize subsections for consistency
PR 7

\subsection{The expression in \eval} \label{sec:minimized}

@@

\subsection{The environments of \eval}\label{sec:env}

@@

\subsection{The origins of \eval}

@@

\subsection{The effects of \eval}

@@

The cases where \eval appears to be overkill include simple expression shapes
such as \k{X}, \k{V}, \k{\$}, \k{<-}, and \k{FUN}. Variable lookup in a remote
environment can be performed by the builtin \k{get} function, a similar case
can be made for assignment and the \k{assign} function. There is a dedicated
\k{do.call} function to build function calls. Unnecessary uses of \eval
often can be uncovered by manual analysis.

\begin{lstlisting}
PerformanceAnalytics::chart.QQPlot <- function (d="norm",dp,...) {
   q.f <- eval(parse(text=paste("q",d)))
   z <- NULL
   eval(parse(text=paste("z<-q.f(",dp,",...)")))
  }
\end{lstlisting}

The above uses \eval to first lookup a target function name from a constructed
string, the it constructs a call to the target that include arguments to the
current functions. \Eval is overkill for this code, a semantically equivalent
function body uses \k{get} to perform the lookup and then calls directly.

\begin{lstlisting}
 function (R, d="norm", dp, ...) {
   q.f <- get(paste0("q",d))
   z <- q.f(dp, ...)
 }
\end{lstlisting}

@@

As a last example, \k{configr::config.funs.par} construct three expressions
that are then fed to \eval. Semantically equivalent code would use \k{get}
to lookup \k{fun}, and then directly execute the three expressions.

@@

We have not yet touched on alternative implementations of \eval. The tidyverse
libraries include their own implementation calls
\k{rlang::eval\_tidy}~\cite{tidyverse}. This departs from \eval in two ways.
First, it supports evaluation of quosures, custom objects used by for
metaprogramming that bundle expressions with an environment. \k{eval\_tidy}
accepts a data mask argument, a set of bindings such as a dataframe or list,
which takes precedence over the environment. Effects such as assignments happen
in the data mask. Expressions such as \k{return()} do not work since the data
mask does not correspond to a frame on the call stack.

>>>>

\subsection{The Expression in \Eval} \label{sec:minimized}

@@

\subsection{The Environments of \Eval}\label{sec:env}

@@

\subsection{The Origins of \Eval}

@@

\subsection{The Effects of \Eval}

@@

The cases where \eval appears to be overkill include simple expression shapes
such as \k{X}, \k{V}, \k{\$}, \k{<-}, and \k{FUN}. Variable lookup in a remote
environment can be performed by the builtin \k{get} function, a similar case
can be made for assignment and the \k{assign} function. There is a dedicated
\k{do.call} function to build function calls. However, the unnecessary uses of
\eval often can only be uncovered by manual analysis.

\begin{lstlisting}[caption={Unnecessary use of \eval},label=lst:unnecessary]
 PerformanceAnalytics::chart.QQPlot <- function (d="norm",dp,...) {
   q.f <- eval(parse(text=paste0("q",d)))
   z <- NULL
   eval(parse(text=paste0("z<-q.f(",dp,",...)")))
\end{lstlisting}

For example in Listing~\ref{lst:unnecessary}, the function uses \eval to lookup
a target function name from a constructed string. Then it constructs a calls it
with the arguments that are passed to the current function. \Eval is overkill
in this case, a semantically equivalent function body uses \k{get} to perform
the lookup and then calls the resulting function directly.

\begin{lstlisting}
 q.f <- get(paste0("q",d))
 z <- q.f(dp, ...)
\end{lstlisting}

@@

% TODO: show how will it look with symbolic field read and write

% TODO: remove - does not bring anything new or interesting
As a last example, \k{configr::config.funs.par} construct three expressions
that are then fed to \eval. Semantically equivalent code would use \k{get}
to lookup \k{fun}, and then directly execute the three expressions.

@@

We have not yet touched on alternative implementations of \eval. The tidyverse
libraries\footnote{An opinionated collection of R packages designed for data
science. It consists of some of the most popular packages.} include their own
implementation calls \k{rlang::eval\_tidy}~\cite{tidyverse}. This departs from
\eval in two ways. First, it supports evaluation of quosures, custom objects
used by for metaprogramming that bundle expressions with an environment.
\k{eval\_tidy} accepts a data mask argument, a set of bindings such as a
dataframe or list, which takes precedence over the environment. Effects such as
assignments happen in the data mask. Expressions such as \k{return()} do not
work since the data mask does not correspond to a frame on the call stack.
